---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
library(ggplot2)
data <- read.csv("./data/activity.csv")
str(data)
data$date <- as.Date(data$date)
summary(data)
```



## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day.  
```{r}
daily_steps <- aggregate(data$steps, list(data$date), sum, na.rm = TRUE)
names(daily_steps) <- c("Date", "Steps")
```
2. Histogram of the total number of steps taken each day .   
```{r}
with(daily_steps, hist(Steps, breaks = seq(0,25000, by=2500), main = "Total number of steps taken per day"))
```  
<br>
3. Calculate and report the mean and median of the total number of steps taken per day.  
```{r}
mean(daily_steps$Steps)
median(daily_steps$Steps)
```



## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).  
``` {r}
avg_daily_steps <- aggregate(data$steps,list(data$interval), mean, na.rm=TRUE)
names(avg_daily_steps) <- c("Interval", "Steps")
with(avg_daily_steps, plot(Interval,Steps, type="l", ylab="Average number of steps"))
```
<br>
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
``` {r}
avg_daily_steps[which.max(avg_daily_steps$Steps),1]
```



## Imputing missing values
1. Calculate and report the total number of missing values in the dataset 
```{r}
sum(is.na(data$steps))
```
2. Devise a strategy for filling in all of the missing values in the dataset.  
Ans. All the missing values are replaced by the mean for that 5-minute interval.
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data1 <- data
for(i in 1:nrow(data1)) {
      if(is.na(data1[i,1])) 
            data1[i,1] <- avg_daily_steps[which(data1[i,3]==avg_daily_steps[,1]),2]
}
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
daily_steps_data1 <- aggregate(data1$steps,list(data1$date), sum, na.rm=TRUE)
names(daily_steps_data1) <- c("Date", "Steps")
with(daily_steps_data1, hist(Steps, breaks = seq(0,25000, by = 2500), main = "Total number of steps taken per day"))
mean(daily_steps_data1$Steps)
median(daily_steps_data1$Steps)
```



## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
data1$day <- weekdays(data1$date)
for(i in 1:nrow(data1)) {
      if(data1[i,4] %in% c("Saturday","Sunday")) data1[i,4] <- "Weekend"
      else data1[i,4] <- "Weekday"
}
data1$day <- as.factor(data1$day)
```
2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken.
```{r}
avg_daily_steps_data1 <- aggregate(data1$steps, list(data1$interval,data1$day), mean, na.rm=TRUE)
names(avg_daily_steps_data1) <- c("Interval", "Day", "Steps")
ggplot(avg_daily_steps_data1, aes(Interval,Steps, col=Day))+
      facet_grid(Day~.)+
      geom_line()+
      ylab("Average number of steps")
```

